---
title: "Competitive FB task - Power analysis - binomial GLMM "
author: "Christoph Völter"
date: "27/01/2021"
output: 
  html_document:
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(cowplot)
library("gghalves")
library(ggthemes)

#load(".RData")
```

## Generate data



```{r echo=FALSE, include=FALSE}
set.seed(1)
n.subject <- 40 # number subjects
n.per.subject <- 24 # observations per subject
n.per.condition <- 24 # observations per subject and condition
n.blocks <- 4
subj.id <- as.factor(paste("subj", str_pad(1:n.subject, 2, pad = "0"), sep = "."))

order_possibilities<-as.factor(c("S-D-E", "S-E-D", "D-S-E", "D-E-S", "E-S-D", "E-D-S"))
first_hiden_food_possibilities <- as.factor(c("Sausage_first", "Dry_food_first"))

locations <- as.vector(replicate(n.subject * n.blocks, sample(x =order_possibilities, size = 6, replace = F)))

first_food<-  as.vector(replicate(n.subject * (n.per.subject/2), sample(x =first_hiden_food_possibilities, size = 2, replace = F)))

start.data <- data.frame(subj.id)
# duplicate rows according to the number obs. per subject:
start.data <- start.data[rep(x = 1:nrow(start.data), times = n.per.subject), ]
start.data <- as.data.frame(start.data)
names(start.data) <- "subj.id"

# add condition and trial number
start.data <- data.frame(expand.grid(subj.id = subj.id, condition = c("preference_test"), trial = c(1:n.per.condition)))%>%
  arrange(subj.id, trial)

start.data$block<-rep((c(rep(1, 6), rep(2, 6), rep(3, 6), rep(4, 6))), n.subject)
start.data$first_food_hidden <- first_food
start.data$locations <- locations

start.data<- start.data%>%
  mutate(loc=locations)%>%
  separate(loc, c("left", "middle", "right"), sep = "-")


```


```{r}
xx <- table(start.data$trial, start.data$locations)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$subj.id, start.data$locations)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$trial, start.data$first_food_hidden)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$subj.id, start.data$first_food_hidden)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$locations, start.data$first_food_hidden)
range(apply(X = xx > 0, MARGIN = 1, sum))

library(summarytools)
view(dfSummary(start.data))
```


```{r echo=FALSE, include=FALSE}
set.seed(1)
n.subject <- 40 # number subjects
n.per.subject <- 24 # observations per subject
n.per.condition <- 12 # observations per subject and condition
subj.id <- as.factor(paste("subj", str_pad(1:n.subject, 2, pad = "0"), sep = "."))
#age_range <- c(0:33) # age range between 7 and 40
#test.per <- c(0.6, 0.7) # performance in occluded condition
#control.per <- c(0.20, 0.30) # performance in visible condition
test.per <- c(0.7) # performance in occluded condition
control.per <- c(0.20) # performance in visible condition


start.data <- data.frame(subj.id)
# duplicate rows according to the number obs. per subject:
start.data <- start.data[rep(x = 1:nrow(start.data), times = n.per.subject), ]
start.data <- as.data.frame(start.data)
names(start.data) <- "subj.id"

# add condition and trial number
start.data <- data.frame(expand.grid(subj.id = subj.id, condition = c("control", "test"), trial = c(1:n.per.condition)))

#add session
start.data$session <- ifelse(start.data$trial<7 & start.data$condition=="control", 1, ifelse(start.data$trial>=7 & start.data$condition=="control", 4, ifelse(start.data$trial<7 & start.data$condition=="test", 2,ifelse(start.data$trial>=7 & start.data$condition=="test", 3, "" ))))

#add trial w/ session
start.data$trial_w_session <- ifelse(start.data$trial<7, start.data$trial, ifelse(start.data$trial>=7, start.data$trial-6, ""))

# z-transformation of covariates
start.data$z.session <- as.vector(scale(as.numeric(start.data$session)))
start.data$z.trial <- as.vector(scale(as.numeric(start.data$trial_w_session )))

# dummy code factors
start.data$condition.test.dummy <- as.numeric(start.data$condition == levels(start.data$condition)[2])


# center condition for random slopes:
start.data$condition.test.c <- start.data$condition.test.dummy - mean(start.data$condition.test.dummy)
```


```{r echo=FALSE, include=FALSE}
# checks:
xx <- table(start.data$subj.id, start.data$condition)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$condition, start.data$session)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$subj.id, start.data$trial_w_session)
range(apply(X = xx > 0, MARGIN = 1, sum))

library(summarytools)
view(dfSummary(start.data))
```


## Simulation

```{r eval=FALSE, include=FALSE}
n.simus <- 1000 # small number for testing
r.effects <- c(1.386294) # random effects to be simulated
# with the intercept being -1.386294 (qlogis(0.2)) we assume a moderately large random intercept of 1.386294.

r.slope.test <- c(2.233592)
# with the estimate being -2.233592(qlogis(0.2)-qlogis(0.7)) we assume a moderately large random slope of 2.233592.

r.slope.session <- 0.2
r.slope.trial <- 0.2
# create object to store the simulation parameters and results:
all.res <- data.frame(expand.grid(
  n.per.subject = n.per.subject, r.effect = r.effects,
  r.slope.test = r.slope.test, r.slope.trial = r.slope.trial, r.slope.session = r.slope.session,
  control.per = control.per,
  test.per = test.per,
  simu = 1:n.simus
))
all.res$icpt <- NA
all.res$conditiontest <- NA
all.res$re.sd <- NA
all.res$warns.full <- NA
all.res$warns.null <- NA
all.res$lrt.p.con <- NA
all.res$lrt.p.age <- NA
all.res$full.null.p <- NA

all.ests <- matrix(NA, nrow = n.simus, ncol = 1)
colnames(all.ests) <- c("lrt.p.con")

# create data frame with design:
## done above

# load packages needed:
library(lme4)
# Loading required package: Matrix
library(kyotil) # we want to store info about convergence issues

# define control structure to make convergence more likely:
contr <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1000000))

xdata <- start.data

# run simulation
for (i in 1:nrow(all.res)) {
  set.seed(i) # allows to later replicate individual simulations


  m.mat <- model.matrix(object = ~ condition + z.session + z.trial, data = xdata) # create model martix

  coefs <- c(
    "(Intercept)" = qlogis(all.res[i, "control.per"]),
    "conditiontest" = log(all.res[i, "test.per"] / (1 - all.res[i, "test.per"])) - log(all.res[i, "control.per"] / (1 - all.res[i, "control.per"])),
    "z.session" = 0,
    "z.trial" = 0
  )

  LP <- m.mat[, names(coefs)] %*% coefs # LP wrt fixed effects

  # add random effect to linear predictor:
  LP <- LP + rnorm(n = n.subject, sd = all.res[i, "r.effect"])[as.numeric(xdata$subj.id)] +
    rnorm(n = n.subject, sd = all.res[i, "r.slope.test"])[as.numeric(xdata$subj.id)] * xdata$condition.test.dummy +
    rnorm(n = n.subject, sd = all.res[i, "r.slope.session"])[as.numeric(xdata$subj.id)] * xdata$z.session +
    rnorm(n = n.subject, sd = all.res[i, "r.slope.trial"])[as.numeric(xdata$subj.id)] * xdata$z.trial

  # generate response:
  xdata$correct <- rbinom(n = nrow(xdata), size = n.per.condition, prob = exp(LP) / (1 + exp(LP)))

  xdata$rv <- cbind(xdata$correct, n.per.condition - xdata$correct)

  # fit full model:
  full <- keepWarnings(glmer(rv ~ condition + z.session + z.trial + (1 + condition.test.c + z.session + z.trial || subj.id),
    data = xdata, family = binomial, control = contr
  ))
  # fit null model:
  null <- keepWarnings(glmer(rv ~ 1 + (1 + condition.test.c + z.session + z.trial || subj.id),
    data = xdata, family = binomial, control = contr
  ))
  # store results:
  all.res[i, c("icpt", "conditiontest", "z.session", "z.trial")] <- fixef(full$value)
  all.res[i, "re.sd"] <- as.data.frame(summary(full$value)$varcor)[1, "sdcor"]
  all.res[i, "warns.full"] <- nchar(paste(full$warnings, collapse = ""))
  all.res[i, "warns.null"] <- nchar(paste(null$warnings, collapse = ""))
  all.res[i, "lrt.p.con"] <- as.data.frame(drop1(full$value, test = "Chisq"))["condition", "Pr(Chi)"]
  all.res[i, "lrt.session.p"] <- as.data.frame(drop1(full$value, test = "Chisq"))["z.session", "Pr(Chi)"]
  all.res[i, "lrt.trial.p"] <- as.data.frame(drop1(full$value, test = "Chisq"))["z.trial", "Pr(Chi)"]
  all.res[i, "full.null.p"] <- as.data.frame(anova(null$value, full$value, test = "Chisq"))[2, "Pr(>Chisq)"]
  print(i)
}

save.image("multiple_possibilities_multiple_possibilities_1000it_6dogs.RData")
```

## Evaluation of results 

* number of warning per combinations of random effects (out of 1000 models per cell)  
Full model:  
```{r echo=FALSE}
#full model
tapply(X=all.res[, "warns.full"]>0, INDEX=all.res[, c("r.slope.test", "r.effect")],
FUN=sum)
#warning codes: 
#363: unable to evaluate scaled gradient. Model failed to converge: degenerate  Hessian with 1 negative eigenvalues
#205: Model is nearly unidentifiable: large eigenvalue ratio - Rescale variables?
```


## Only models that converged are evaluated from here on:  

```{r include=FALSE}
all.res2=subset(all.res, warns.full==0)
```


### How many models converged, have a significant full-null model comparison, and a significant LRT of condition?  
```{r echo=FALSE}
n.converged<- all.res2%>%
    group_by(test.per,r.effect, r.slope.test ) %>%
  summarise(n.converged=length(lrt.p.con))


lrt.data2 <- all.res2 %>%
  #filter(full.null.p<0.05)%>%
  group_by(test.per, control.per, r.effect, r.slope.test ) %>%
  summarise(lrt.p.con.median = median(lrt.p.con), 
            lrt.p.session.median = median(lrt.session.p),
            n.sign.lrt.con = length(lrt.p.con[lrt.p.con < 0.05]),
            n.sign.lrt.session = length(lrt.session.p[lrt.session.p < 0.05]),
            n.lrt = n.simus,
            proportion.sign.lrt.con = length(lrt.p.con[lrt.p.con < 0.05]) / n.simus,
            proportion.sign.lrt.session = length(lrt.session.p[lrt.session.p < 0.05]) / n.simus)%>%
  full_join(n.converged)

lrt.data2
```

#### Plotting the proportion of significant LRTs for the predictor variable condition ONLY based on models that converged and with a significant full-null model comparison

```{r echo=FALSE}
p.con.power <- ggplot(data = lrt.data2, aes(x= as.factor(test.per),y = proportion.sign.lrt.con, fill=as.factor(control.per))) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1))+
  geom_hline(yintercept = 0.8, colour = "black", lwd = 1.1, lty = 2) +
   # geom_hline(yintercept = 0.05, colour = "darkgrey", lwd = 1.1, lty = 4) +
  scale_fill_manual(values=c("dodgerblue", "darkorange"))+
  labs(fill = "Control condition", y="Power", x= "Test condition") +
  theme_few()#+
  #theme(legend.position="none")
p.con.power

#ggsave(p.con.power, filename = "graphics/ExplSeek_power_glmm_cbind_30sub.png", scale = 0.8, height = 5, width = 5)
```


#### Plotting the fixed effect of condition

```{r echo=FALSE}
p.con <- ggplot(data = all.res2, aes(x= as.factor(test.per), fill=as.factor(control.per)))  +
  geom_jitter(data = all.res2, aes(x = as.factor(test.per), y = conditiontest, color = as.factor(control.per)), size = 1.5, position = position_jitterdodge(dodge.width = 0.8, jitter.width = 0.5), alpha = .1) +
  scale_color_manual(values = c("dodgerblue", "darkorange")) +#, name = "control condition"
  scale_fill_manual(values = c("dodgerblue", "darkorange")) +#, name = "control condition"
    geom_boxplot(data = all.res2 %>% filter(control.per == "0.2"), aes(x = as.factor(test.per), y = conditiontest), position = position_nudge(x = -.2), width = 0.3, alpha = 0.5, outlier.colour = "white") +
  geom_boxplot(data = all.res2 %>% filter(control.per == "0.3"), aes(x = as.factor(test.per), y = conditiontest), position = position_nudge(x = .2), width = 0.3, alpha = 0.5, outlier.colour = "white") +
  #geom_hline(data = data.frame(test.per = "0.7"), aes(yintercept = coefs["conditiontest"]), colour = "black", lwd = 1.1, lty = 2, alpha = 0.7) +
  geom_segment( aes(x=2, xend=2.4, y = qlogis(0.7) - qlogis(0.3), yend = qlogis(0.7) - qlogis(0.3)), colour = "red", lwd = 1, lty = 4, alpha = 0.7) +
    geom_segment( aes(x=1.6, xend=2, y = qlogis(0.7) - qlogis(0.2), yend = qlogis(0.7) - qlogis(0.2)), colour = "red", lwd = 1, lty = 2, alpha = 0.7) +
    geom_segment( aes(x=0.6, xend=1, y = qlogis(0.6) - qlogis(0.2), yend = qlogis(0.6) - qlogis(0.2) ), colour = "red", lwd = 1, lty = 2, alpha = 0.7) +
    geom_segment( aes(x=1, xend=1.4, y = qlogis(0.6) - qlogis(0.3), yend = qlogis(0.6) - qlogis(0.3) ), colour = "red", lwd = 1, lty = 4, alpha = 0.7) +
  ylab("Condition (fixed effect)") +
  xlab("Test condition") +
  theme_few() +
  #theme(legend.title = element_blank())+
  theme(legend.position = "none")

p.con
```

```{r}

pg<-plot_grid(p.con, p.con.power, nrow=1, labels=c("A", "B"))
ggsave(pg, filename = "graphs/simulation_estimates_power.png", width=10, height=5, scale=0.7)
```

```{r}
save.image("multiple_possibilities_multiple_possibilities_1000it_6dogs.RData")
```


